### 2021-11-15 chapter2/chapter.tex
nigel changes ch 3 next set 1, update QBDRAP and fluid queue to use vector notation, introduce notation for tau_n<t<=tau_n+1

%Let \(Z\sim ME_p(\bs \alpha, \bs{S}, \bs s)\) of dimension \(p\) with \(\bs\alpha\bs e = 1\), \(\bs s = -\bs{S}\bs e\). For each \(i\in\mathcal S\), define \(\bs{S}_i = |c_i|\bs{S}\) and \(\bs s_i = |c_i|\bs s\). Let \(\{\bs A(t)\} = \left\{\bs A(t)\right\}_{t\geq 0}\) be a \(p\)-dimensional vector-valued stochastic process which is piecewise-deterministic and c\`adl\`ag. Additionally, let \(\left\{{\phi}(t)\right\}_{t\geq 0}\) be a real-valued stochastic process. Together, \(\left\{(\bs A(t),{\phi}(t))\right\}_{t\geq 0}\), form a Markov process, however, \(\{\bs A(t)\}\) alone is not Markovian. We will show later that \(\left\{{\phi}(t)\right\}\) is a CTMC the same as \(\left\{\varphi(t)\right\}\). The sample path behaviour of \(\{\bs A(t)\}\) and \({\phi}(t)\) is as follows. 
\section{The dynamics of the QBD-RAP approximation}\label{sec: dynamics}
We now have all the elements we need to describe the dynamics of the QBD-RAP approximation. Since the phase dynamics are a CTMC, we choose to use an alternate notation to the standard presented in Section \ref{sec: qbd-rap}. Let \(\{(L(t),\bs A(t),\phi(t))\}_{t\geq0}\) be the QBD-RAP approximation of a fluid queue, where \(\{L(t)\}\) is the level, \(\{\bs A(t)\}\) is the orbit process and \(\{\phi(t)\}\) is the phase process. We use this representation to make explicit how the approximation captures the phase dynamics of the fluid queue. We will show later that \(\phi(t)\) captures the phase dynamics of the fluid queue exactly, provided the phase process is independent of the fluid level \(\{X(t)\}\). The level \(L(t)=\ell\) approximates which band \(\calD_\ell\) that \(X(t)\) is in at time \(t\), and the orbit \(\bs A(t)\) can be use to obtain an approximation of where \(X(t)\) is within the interval \(\calD_\ell\).

@@

%To motivate the construction, first consider the initial condition for the QBD-RAP approximation, \((L(0),\bs A(0), \varphi(0))=(\ell,\bs \alpha,i)\), corresponding to the initial condition of the fluid queue \((X(0),\varphi(0))=(y_\ell,i),\) for some \(i\in\mathcal S_+\). After time \(t\), \(c_it\in[0,\Delta)\), and assuming that no events of the QBD-RAP have occurred by this time, the position of the QBD-RAP is \((L(t),\bs A(t),\varphi(t)) = (\ell, \bs A(t), i),\) where \(\bs A(t) = \cfrac{\bs \alpha e^{c_i\bs{S}t}}{\bs \alpha e^{c_i\bs{S}t}\bs e}\). The level of the fluid at the same time is \(X(t) = y_\ell + c_i t\). 
We argued earlier that we can think of the orbit \(\bs k(x)\) as corresponding to the fluid being a distance of \(x\) from the left boundary of an interval when \(i\in\calS_+\), or from the right boundary of an interval when \(i\in\calS_-\). With this interpretation, an approximation to the initial condition \((X(0),\varphi(0))=(x_0,i),\,x_0\in\mathcal D_{\ell,i},\,i\in\mathcal S_+\cup\calS_{+0}\) is to set the initial orbit to \(\bs k(x_0-y_{\ell})\). Similarly, an approximation to the initial condition \((X(0),\varphi(0))=(x_0,i), \,x_0\in\mathcal D_{\ell,i}, x_0\neq y_{\ell+1},\,i\in\mathcal S_-\cup\calS_{-0}\) is to set the initial orbit position to \(\bs k(y_{\ell+1}-x_0).\)

@@

A solution to this problem is to augment a set of ephemeral states to the QBD-RAP which capture the initial sojourn in \(\calS_0\) only. Let us denote this set of ephemeral states \(\calS_0^*\) which has the same phase dynamics as \(\calS_0\). Suppose we wish to approximate a fluid queue with the initial condition \(\mu_k\) where \(\mu_k\) has mass 1. At time \(0\), we start the QBD-RAP in the ephemeral set of states \(\calS_0^*\) in phase \(k^*\in\calS_0^*\). While in \(\calS_0^*\), the phase process evolves according to the generator \(\bs T_{00}\). Upon exiting the ephemeral set \(\calS_0^*\) at time \(t\), say, the QBD-RAP process jumps to \(L(t)=\ell,\, \bs A(t)=\int_{\calD_{\ell,i}}\bs a_{\ell,i}(x)\wrt \mu_k,\,\varphi(t)=i\), where \(\ell\in\mathcal K\), \(i\in\calS_+\cup\calS_-\).

Notice that we have not mentioned anything about the level or orbit process during the sojourn in \(\calS_0^*\). This is because all information about the level and phase during this time is contained within the initial condition \(\mu_k\), and the only thing that changes is the phase.

To include the ephemeral states within the generator of the QBD-RAP we can do the following, 
\begin{align*}
	&\left[\begin{array}{c|cccc}
		\bs T_{00} & \bs B_{-1}^* &\bs B_{0}^*&\bs B_{1}^* & \hdots \\\hline
		&\check{\bs B}_{0} &  \check{\bs B}_{+1} &&\\
		& \check{\bs B}_{-1}&\check{\bs B}_{0}&\bs B_{+1}& \\
		&&\bs B_{-1} & \bs B_0&\ddots \\
		&&&\ddots&\ddots 
	\end{array}\right],
\end{align*}
where 
\begin{align*}
	\bs B_{-1}^* &= \vligne{\bs T_{0-}\mu_k(\{0\}) & \bs 0} & \bs B_{k}^* &= \vligne{\bs T_{0+}\otimes \bs a_k^+ & \bs 0 & \bs T_{0-}\otimes \bs a_k^- & \bs 0},
	%
	\\\check{\bs B}_{0} &= \left[\begin{array}{cc}\bs T_{--} & \bs T_{-0}\\\bs T_{0-} & \bs T_{00}\end{array}\right], 
	& \check{\bs B}_{+1} &= \left[\begin{array}{cccc}\bs T_{-+}\otimes \bs \alpha & \bs 0 & \bs 0 & \bs 0 \\ \bs T_{0+}\otimes \bs \alpha & \bs 0 & \bs 0 & \bs 0 \end{array}\right],
	%
	\\\check{\bs B}_{-1} &= \left[\begin{array}{cc} \bs0&\bs0\\\bs0&\bs0\\(\bs C_-\bs P_{--})\otimes \bs s & (\bs C_-\bs P_{-0})\otimes \bs s \\\bs0&\bs0\end{array}\right], &&
\end{align*}
and 
\begin{align*}
	\bs a_\ell^r = \int_{\calD_{\ell,r}} \bs a_{\ell,r}(x)\wrt \mu_k,\, \ell\in\mathcal K\setminus\{-1,K+1\},\, r\in\{+,-\}. 
\end{align*}

@@

This reasoning leads to an approximation of the distribution of the fluid at time \(t\), \(\mathbb P(X(t)\in \wrt x , \varphi(t) = j\mid X(0)=x_0, \varphi(0) = i) \), as
\begin{align}
		&\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs a_{\ell_0,i}(x_0), \phi(0) = i)  
		\bs a e^{\bs{S}z} \bs s \wrt x,\label{eqn: density approx aug}
		% \intertext{for \(j\in\mathcal S_+\cup\calS_{+0}\) and}
		% &\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs a_{\ell_0,i}(x_0), \phi(0) = i) \nonumber 
		% \\&\qquad\times\bs a e^{\bs{S}(x-y_\ell)} \bs s \wrt x,\label{eqn: density approx aug2}
\end{align}
where \(x\in\mathcal D_{\ell,j}\), \(x_0\in\mathcal D_{\ell_0,i}\) and 
\[
	z = \begin{cases} 
		x - y_\ell & i\in\calS_- \\
		y_{\ell+1} - x & i\in\calS_+.
	\end{cases}
\]
%For example, for the augmented scheme in Section~\ref{subsec: augmented }, we approximate \(\mathbb P(X(t)\in \wrt x , \varphi(t) = j\mid X(0)=x_0, \varphi(0) = i)\) by 
%\begin{align}
%	\label{eqn: density approx aug}
%		&\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\varphi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs a, \varphi(0) = i) \bs a e^{\bs{S}(y_{\ell+1}-x)} \bs s \wrt x & j\in\mathcal S_+\cup \mathcal S_{+0},
%		\\\label{eqn: density approx aug2}
%		&\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\varphi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs a, \varphi(0) = i) \bs a e^{\bs{S}(x-y_\ell)} \bs s \wrt x & j\in\mathcal S_-\cup \mathcal S_{-0},
%\end{align}

While the estimate (\ref{eqn: density approx aug}) is appealing, it is, however, defective -- it does not integrate to 1 for any finite-dimensional matrix exponential distribution. To see this, compute 
\begin{align*}
	&\sum_{\ell\in\mathcal K}\sum_{j\in\mathcal S}\int_{z\in\mathcal D_{\ell,j}}\displaystyle\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a, \phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs  a_{\ell,i}(x_0), 
	\\&\qquad{}\qquad{}\phi(0) = i) \bs a e^{ \bs{S}z} \bs s \wrt z
	% \\&\qquad{}{}+\sum_{\ell\in\mathcal K}\sum_{j\in\mathcal S_-\cup \calS_{-0}}\int_{x\in\mathcal D_{\ell,j}}\displaystyle\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a, \phi(t) = j\mid L(0)=\ell_0, 
	% \\&\qquad{}\qquad{}\bs A(0) = \bs  a_{\ell,i}(x_0), \phi(0) = i) \bs a e^{ \bs{S}(x-y_\ell)} \bs s \wrt x
	\\&= \sum_{\ell\in\mathcal K}\sum_{j\in\mathcal S}\displaystyle\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a, \phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs  a_{\ell,i}(x_0), 
	\\&\qquad{}\qquad{}\phi(0) = i) (1-\bs a e^{ \bs{S}\Delta}\bs e  )
	\\&<1,
\end{align*}
since \(\bs a e^{\bs{S}u} \bs e>0\) for any \(u>0\). For an orbit position \(\bs A(t)=\bs a\), the amount of mass missing from each level and phase is 
\[\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs  a_{\ell,i}(x_0), \phi(0) = i) \bs a e^{\bs{S}\Delta} \bs e \wrt x.\]
The problem is that the triple \((\bs a,\bs S, \bs s)\) defines a distribution on \(x\in[0,\infty)\), however, it only makes sense to take \(x\in[0,\Delta)\) in the approximation scheme.

@@

To partially rectify this we can instead approximate \(\mathbb P(X(t)\in \wrt x  , \varphi(t) = j\mid X(0)=x_0, \varphi(0) = i) \) by
\begin{align}
%	&\mathbb P(X(t)\in \wrt x  , \varphi(t) = j\mid X(0)=x_0, \varphi(0) = i) \nonumber
	\label{eqn: density approx 3+}&
		\displaystyle\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs  a_{\ell,i}(x_0), \phi(0) = i) \bs a u_i^\ell(x) \wrt x ,
\end{align}
{where}
\begin{align}
		\bs a u_i^\ell(x) = \begin{cases}
			\bs a \left(e^{ \bs{S}(y_{\ell+1}-x)} \bs s + e^{ \bs{S}(2\Delta-(y_{\ell+1}-x))} \bs s\right), & i\in\calS_+, \\ 
			\bs a \left(e^{ \bs{S}(x-y_\ell)} \bs s+e^{ \bs{S}(2\Delta-(x-y_\ell))} \bs s\right), & i\in\calS_-. \\ 
		\end{cases}\label{eqn: density approx 3+a}
\end{align}
Intuitively, we take the density function \(\bs ae^{\bs{S}x}\bs s\) on \(x \in [0,2\Delta)\), and `fold' it back on itself around \(\Delta\), to create a density function on \([0,\Delta)\), 
\[\bs ae^{\bs{S}x}\bs s +\bs ae^{\bs{S}(2\Delta-x)}\bs s,\, x \in [0,\Delta).\]
The missing mass is now proportional to 
\[\bs a e^{\bs S 2\Delta}\bs e,\]
which is less than the quantity \(\bs a e^{\bs S \Delta}\bs e\) from the approximation scheme (\ref{eqn: density approx aug}). There is nothing special about the choice of truncating the distribution at \(2\Delta\) in this construction, and we could choose any truncation value greater than \(\Delta\).

A third option is to approximate \(\mathbb P(X(t)\in \wrt x  , \varphi(t) = j\mid X(0)=x_0, \varphi(0) = i) \) by
\begin{align}
%	&\mathbb P(X(t)\in \wrt x  , \varphi(t) = j\mid X(0)=x_0, \varphi(0) = i) \nonumber
	\label{eqn: density approx 4+}&
		\displaystyle\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs  a_{\ell,i}(x_0), \phi(0) = i) \bs a v_i^\ell(x) \wrt x ,
\end{align}
{where}
\begin{align}
		\bs a v_i^\ell(x) = \begin{cases}
			\bs a \left(e^{ \bs{S}(y_{\ell+1}-x)} + e^{ \bs{S}(2\Delta-(y_{\ell+1}-x))} \right)\left[I-e^{\bs S 2\Delta}\right]^{-1}\bs s, & i\in\calS_+, \\ 
			\bs a \left(e^{ \bs{S}(x-y_\ell)}+e^{ \bs{S}(2\Delta-(x-y_\ell))}\right) \left[I-e^{\bs S 2\Delta}\right]^{-1}\bs s, & i\in\calS_-. \\ 
		\end{cases}
\end{align}
Here, we take the density function \(\bs a e^{\bs S x}\bs s\), defined on \([0,\infty)\), and map it to a density function on \([0,\Delta)\) by \(\Delta - |(x \mod 2\Delta )-\Delta|\). The resulting density function is 
\begin{align*}
	v(x) &= \sum_{m=0}^\infty \bs a \left(e^{\bs S (x+2\Delta m)} + e^{\bs S (2\Delta -x+2\Delta m)} \right)\bs s
	%
	\\&= \bs a \left(e^{\bs S x} + e^{\bs S (2\Delta -x)} \right)\sum_{m=0}^\infty e^{\bs S 2\Delta m} \bs s
	%
	\\&= \bs a \left(e^{\bs S x} + e^{\bs S (2\Delta -x)} \right)\left[I-e^{S2\Delta}\right]^{-1} \bs s.
\end{align*}

>>>>

%Let \(Z\sim ME_p(\bs \alpha, \bs{S}, \bs s)\) of dimension \(p\) with \(\bs\alpha\bs e = 1\), \(\bs s = -\bs{S}\bs e\). For each \(i\in\mathcal S\), define \(\bs{S}_i = |c_i|\bs{S}\) and \(\bs s_i = |c_i|\bs s\). Let \(\{\bs A(t)\} = \left\{\bs A(t)\right\}_{t\geq 0}\) be a \(p\)-dimensional vector-valued stochastic process which is piecewise-deterministic and c\`adl\`ag. Additionally, let \(\left\{{\phi}(t)\right\}_{t\geq 0}\) be a real-valued stochastic process. Together, \(\left\{(\bs A(t),{\phi}(t))\right\}_{t\geq 0}\), form a Markov process, however, \(\{\bs A(t)\}\) alone is not Markovian. We will show later that \(\left\{{\phi}(t)\right\}\) is a CTMC the same as \(\left\{\varphi(t)\right\}\). The sample path behaviour of \(\{\bs A(t)\}\) and \({\phi}(t)\) is as follows. 
\section{The dynamics of the QBD-RAP approximation}\label{sec: dynamics}
We now have all the elements we need to describe the dynamics of the QBD-RAP approximation. Since the phase dynamics are a CTMC, we choose to use an alternate notation to the standard presented in Section \ref{sec: qbd-rap}. Let \(\{\bs Y(t)\}_{t\geq0} = \{(L(t),\bs A(t),\phi(t))\}_{t\geq0}\) be the QBD-RAP approximation of a fluid queue, where \(\{L(t)\}\) is the level, \(\{\bs A(t)\}\) is the orbit process and \(\{\phi(t)\}\) is the phase process. We use this representation to make explicit how the approximation captures the phase dynamics of the fluid queue. Furthermore, define the notation \(\bs Y_d(t) = (L(t),\phi(t))\) for the discrete elements of \(\bs Y(t)\).

We will show later that \(\phi(t)\) captures the phase dynamics of the fluid queue exactly, provided the phase process is independent of the fluid level \(\{X(t)\}\). The level \(L(t)=\ell\) approximates which band \(\calD_\ell\) that \(X(t)\) is in at time \(t\), and the orbit \(\bs A(t)\) can be use to obtain an approximation of where \(X(t)\) is within the interval \(\calD_\ell\).

@@

%To motivate the construction, first consider the initial condition for the QBD-RAP approximation, \((L(0),\bs A(0), \varphi(0))=(\ell,\bs \alpha,i)\), corresponding to the initial condition of the fluid queue \((X(0),\varphi(0))=(y_\ell,i),\) for some \(i\in\mathcal S_+\). After time \(t\), \(c_it\in[0,\Delta)\), and assuming that no events of the QBD-RAP have occurred by this time, the position of the QBD-RAP is \((L(t),\bs A(t),\varphi(t)) = (\ell, \bs A(t), i),\) where \(\bs A(t) = \cfrac{\bs \alpha e^{c_i\bs{S}t}}{\bs \alpha e^{c_i\bs{S}t}\bs e}\). The level of the fluid at the same time is \(X(t) = y_\ell + c_i t\). 
We argued earlier that we can think of the orbit \(\bs k(x)\) as corresponding to the fluid being a distance of \(x\) from the left boundary of an interval when \(i\in\calS_+\), or from the right boundary of an interval when \(i\in\calS_-\). With this interpretation, an approximation to the initial condition \(\bs X(0) = (X(0),\varphi(0))=(x_0,i),\,x_0\in\mathcal D_{\ell,i},\,i\in\mathcal S_+\cup\calS_{+0}\) is to set the initial orbit to \(\bs k(x_0-y_{\ell})\). Similarly, an approximation to the initial condition \(\bs X(0)=(x_0,i), \,x_0\in\mathcal D_{\ell,i}, x_0\neq y_{\ell+1},\,i\in\mathcal S_-\cup\calS_{-0}\) is to set the initial orbit position to \(\bs k(y_{\ell+1}-x_0).\)

@@

A solution to this problem is to augment a set of ephemeral states to the QBD-RAP which capture the initial sojourn in \(\calS_0\) only. Suppose we wish to approximate a fluid queue with the initial condition \(\mu_k\) where \(\mu_k\) has mass 1. Let us denote the set of ephemeral states by \(\calS_0^{*,k}\) and each state in \(\calS_0^{*,k}\) by \(i^*\). Each state \(i^*\in\calS_0^{*,k}\) is a copy of \(i\in\calS_0\), and, in \(\calS_0^{*,k}\) the phase process has the same phase dynamics as it does in \(\calS_0\). At time \(0\), we start the QBD-RAP in the ephemeral set of states \(\calS_0^{*,k}\) in phase \(k^*\in\calS_0^{*,k}\). While in \(\calS_0^{*,k}\), the phase process evolves according to the generator \(\bs T_{00}\). Upon exiting the ephemeral set \(\calS_0^{*,k}\) at time \(t\), say, the QBD-RAP process jumps to \(L(t)=\ell,\, \bs A(t)=\int_{\calD_{\ell,i}}\bs a_{\ell,i}(x)\wrt \mu_k,\,\varphi(t)=i\), where \(\ell\in\mathcal K\), \(i\in\calS_+\cup\calS_-\).

Notice that we have not mentioned anything about the level or orbit process during the sojourn in \(\calS_0^{*,k}\). This is because all information about the level and phase during this time is contained within the initial condition \(\mu_k\), and the only thing that changes is the phase.

To include the ephemeral states within the generator of the QBD-RAP we can do the following, 
\begin{align*}
	&\left[\begin{array}{c|cccc}
		\bs T_{00} & \bs B_{-1}^{*,k} &\bs B_{0}^{*,k}&\bs B_{1}^{*,k} & \hdots \\\hline
		&\check{\bs B}_{0} &  \check{\bs B}_{+1} &&\\
		& \check{\bs B}_{-1}&\check{\bs B}_{0}&\bs B_{+1}& \\
		&&\bs B_{-1} & \bs B_0&\ddots \\
		&&&\ddots&\ddots 
	\end{array}\right],
\end{align*}
where 
\begin{align*}
	\bs B_{-1}^{*,k} &= \vligne{\bs T_{0-}\mu_k(\{0\}) & \bs 0} & \bs B_{\ell}^{*,k} &= \vligne{\bs T_{0+}\otimes \bs a_\ell^+ & \bs 0 & \bs T_{0-}\otimes \bs a_\ell^- & \bs 0},
	%
	\\\check{\bs B}_{0} &= \left[\begin{array}{cc}\bs T_{--} & \bs T_{-0}\\\bs T_{0-} & \bs T_{00}\end{array}\right], 
	& \check{\bs B}_{+1} &= \left[\begin{array}{cccc}\bs T_{-+}\otimes \bs \alpha & \bs 0 & \bs 0 & \bs 0 \\ \bs T_{0+}\otimes \bs \alpha & \bs 0 & \bs 0 & \bs 0 \end{array}\right],
	%
	\\\check{\bs B}_{-1} &= \left[\begin{array}{cc} \bs0&\bs0\\\bs0&\bs0\\(\bs C_-\bs P_{--})\otimes \bs s & (\bs C_-\bs P_{-0})\otimes \bs s \\\bs0&\bs0\end{array}\right], &&
\end{align*}
and 
\begin{align*}
	\bs a_\ell^r = \int_{\calD_{\ell,r}} \bs a_{\ell,r}(x)\wrt \mu_k,\, \ell\in\mathcal K\setminus\{-1,K+1\},\, r\in\{+,-\}. 
\end{align*}

In general, we need to augment a set of states \(\calS_0^{*,k}\) to the QBD-RAP for each initial phase, \(k\in\calS_0\) such that \(\mu_k\) has positive mass.

@@

This reasoning leads to an approximation of the distribution of the fluid at time \(t\), \(\mathbb P(\bs X(t)\in (\wrt x,j) \mid \bs X(0)=(x_0,i)) \), as
\begin{align}
		&\int_{\bs a \in\mathcal A}\mathbb P(\bs Y(t) \in (\ell, \wrt \bs a, j)\mid \bs Y(0)=(\ell_0, \bs a_{\ell_0,i}(x_0), i) )
		\bs a e^{\bs{S}z} \bs s \wrt x,
		%
		% \\&\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs a_{\ell_0,i}(x_0), \phi(0) = i)  
		% \bs a e^{\bs{S}z} \bs s \wrt x,
		\label{eqn: density approx aug}
		% \intertext{for \(j\in\mathcal S_+\cup\calS_{+0}\) and}
		% &\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs a_{\ell_0,i}(x_0), \phi(0) = i) \nonumber 
		% \\&\qquad\times\bs a e^{\bs{S}(x-y_\ell)} \bs s \wrt x,\label{eqn: density approx aug2}
\end{align}
where \(x\in\mathcal D_{\ell,j}\), \(x_0\in\mathcal D_{\ell_0,i}\) and 
\[
	z = \begin{cases} 
		x - y_\ell & i\in\calS_-, \\
		y_{\ell+1} - x & i\in\calS_+.
	\end{cases}
\]
%For example, for the augmented scheme in Section~\ref{subsec: augmented }, we approximate \(\mathbb P(X(t)\in \wrt x , \varphi(t) = j\mid X(0)=x_0, \varphi(0) = i)\) by 
%\begin{align}
%	\label{eqn: density approx aug}
%		&\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\varphi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs a, \varphi(0) = i) \bs a e^{\bs{S}(y_{\ell+1}-x)} \bs s \wrt x & j\in\mathcal S_+\cup \mathcal S_{+0},
%		\\\label{eqn: density approx aug2}
%		&\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\varphi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs a, \varphi(0) = i) \bs a e^{\bs{S}(x-y_\ell)} \bs s \wrt x & j\in\mathcal S_-\cup \mathcal S_{-0},
%\end{align}

While the estimate (\ref{eqn: density approx aug}) is appealing, it is, however, defective -- it does not integrate to 1 for any finite-dimensional matrix exponential distribution. To see this, compute 
\begin{align*}
	&\sum_{\ell\in\mathcal K}\sum_{j\in\mathcal S}\int_{z\in\mathcal D_{\ell,j}}\displaystyle\int_{\bs a \in\mathcal A}\mathbb P(\bs Y(t) \in (\ell, \wrt \bs a, j)\mid \bs Y(0)=(\ell_0, \bs  a_{\ell,i}(x_0), i)) \bs a e^{ \bs{S}z} \bs s \wrt z
	% \\&\qquad{}{}+\sum_{\ell\in\mathcal K}\sum_{j\in\mathcal S_-\cup \calS_{-0}}\int_{x\in\mathcal D_{\ell,j}}\displaystyle\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a, \phi(t) = j\mid L(0)=\ell_0, 
	% \\&\qquad{}\qquad{}\bs A(0) = \bs  a_{\ell,i}(x_0), \phi(0) = i) \bs a e^{ \bs{S}(x-y_\ell)} \bs s \wrt x
	\\&= \sum_{\ell\in\mathcal K}\sum_{j\in\mathcal S}\displaystyle\int_{\bs a \in\mathcal A}\mathbb P(\bs Y(t) \in (\ell, \wrt \bs a, j)\mid \bs Y(\ell_0, \bs  a_{\ell,i}(x_0), i)) (1-\bs a e^{ \bs{S}\Delta}\bs e  )
	\\&<1,
	% &\sum_{\ell\in\mathcal K}\sum_{j\in\mathcal S}\int_{z\in\mathcal D_{\ell,j}}\displaystyle\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a, \phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs  a_{\ell,i}(x_0), 
	% \\&\qquad{}\qquad{}\phi(0) = i) \bs a e^{ \bs{S}z} \bs s \wrt z
	% % \\&\qquad{}{}+\sum_{\ell\in\mathcal K}\sum_{j\in\mathcal S_-\cup \calS_{-0}}\int_{x\in\mathcal D_{\ell,j}}\displaystyle\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a, \phi(t) = j\mid L(0)=\ell_0, 
	% % \\&\qquad{}\qquad{}\bs A(0) = \bs  a_{\ell,i}(x_0), \phi(0) = i) \bs a e^{ \bs{S}(x-y_\ell)} \bs s \wrt x
	% \\&= \sum_{\ell\in\mathcal K}\sum_{j\in\mathcal S}\displaystyle\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a, \phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs  a_{\ell,i}(x_0), 
	% \\&\qquad{}\qquad{}\phi(0) = i) (1-\bs a e^{ \bs{S}\Delta}\bs e  )
	% \\&<1,
\end{align*}
since \(\bs a e^{\bs{S}u} \bs e>0\) for any \(u>0\). For an orbit position \(\bs A(t)=\bs a\), the amount of mass missing from each level and phase is 
\[\mathbb P(\bs Y(t)\in (\ell, \wrt \bs a,j)\mid \bs Y(0)=(\ell_0,  \bs  a_{\ell,i}(x_0), i)) \bs a e^{\bs{S}\Delta} \bs e \wrt x.\]
The problem is that the triple \((\bs a,\bs S, \bs s)\) defines a distribution on \(x\in[0,\infty)\), however, it only makes sense to take \(x\in[0,\Delta)\) in the approximation scheme.

@@

To partially rectify this we can instead approximate \(\mathbb P(\bs X(t)\in (\wrt x, j)\mid \bs X(0)=(x_0, i)) \) by
\begin{align}
%	&\mathbb P(X(t)\in \wrt x  , \varphi(t) = j\mid X(0)=x_0, \varphi(0) = i) \nonumber
	\label{eqn: density approx 3+}&
	\displaystyle\int_{\bs a \in\mathcal A}\mathbb P(\bs Y(t) \in(\ell, \wrt \bs a, j)\mid \bs Y(0)=()\ell_0, \bs  a_{\ell,i}(x_0), i)) \bs a u_i^\ell(x) \wrt x ,
	% \displaystyle\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs  a_{\ell,i}(x_0), \phi(0) = i) \bs a u_i^\ell(x) \wrt x ,
\end{align}
{where}
\begin{align}
		\bs a u_i^\ell(x) = \begin{cases}
			\bs a \left(e^{ \bs{S}(y_{\ell+1}-x)} \bs s + e^{ \bs{S}(2\Delta-(y_{\ell+1}-x))} \bs s\right), & i\in\calS_+, \\ 
			\bs a \left(e^{ \bs{S}(x-y_\ell)} \bs s+e^{ \bs{S}(2\Delta-(x-y_\ell))} \bs s\right), & i\in\calS_-. \\ 
		\end{cases}\label{eqn: density approx 3+a}
\end{align}
Intuitively, we take the density function \(\bs ae^{\bs{S}x}\bs s\) on \(x \in [0,2\Delta)\), and `fold' it back on itself around \(\Delta\), to create a density function on \([0,\Delta)\), 
\[\bs ae^{\bs{S}x}\bs s +\bs ae^{\bs{S}(2\Delta-x)}\bs s,\, x \in [0,\Delta).\]
The missing mass is now proportional to 
\[\bs a e^{\bs S 2\Delta}\bs e,\]
which is less than the quantity \(\bs a e^{\bs S \Delta}\bs e\) from the approximation scheme (\ref{eqn: density approx aug}). There is nothing special about the choice of truncating the distribution at \(2\Delta\) in this construction, and we could choose any truncation value greater than \(\Delta\).

A third option is to approximate \(\mathbb P(\bs X(t)\in (\wrt x, j)\mid \bs X(0)=(x_0, i)) \) by
\begin{align}
%	&\mathbb P(X(t)\in \wrt x  , \varphi(t) = j\mid X(0)=x_0, \varphi(0) = i) \nonumber
	\label{eqn: density approx 4+}&
		\displaystyle\int_{\bs a \in\mathcal A}\mathbb P(\bs Y(t) \in (\ell, \wrt \bs a,j)\mid \bs Y(0) = (\ell_0, \bs  a_{\ell,i}(x_0), i)) \bs a v_i^\ell(x) \wrt x ,
\end{align}
{where}
\begin{align}
		\bs a v_i^\ell(x) = \begin{cases}
			\bs a \left(e^{ \bs{S}(y_{\ell+1}-x)} + e^{ \bs{S}(2\Delta-(y_{\ell+1}-x))} \right)\left[I-e^{\bs S 2\Delta}\right]^{-1}\bs s, & i\in\calS_+, \\ 
			\bs a \left(e^{ \bs{S}(x-y_\ell)}+e^{ \bs{S}(2\Delta-(x-y_\ell))}\right) \left[I-e^{\bs S 2\Delta}\right]^{-1}\bs s, & i\in\calS_-. \\ 
		\end{cases}
\end{align}
Here, we take the density function \(\bs a e^{\bs S x}\bs s\), defined on \([0,\infty)\), and map it to a density function on \([0,\Delta)\) by \(\Delta - |(x \mod 2\Delta )-\Delta|\). The resulting density function is 
\begin{align*}
	v(x) &= \sum_{m=0}^\infty \bs a \left(e^{\bs S (x+2\Delta m)} + e^{\bs S (2\Delta -x+2\Delta m)} \right)\bs s
	%
	\\&= \bs a \left(e^{\bs S x} + e^{\bs S (2\Delta -x)} \right)\sum_{m=0}^\infty e^{\bs S 2\Delta m} \bs s
	%
	\\&= \bs a \left(e^{\bs S x} + e^{\bs S (2\Delta -x)} \right)\left[I-e^{S2\Delta}\right]^{-1} \bs s.
\end{align*}
